---
title: "Residuals Explanation"
author: "Brandon Ritchie"
date: "9/20/2021"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(grid)
library(pBrackets)
# Home
read_csv("C:/Users/brand/Documents/Statistics-Notebook/Data/laptops.csv")
# Campus
#data <- read_csv("C:/Users/britchi3/Documents/Statistics-Notebook/Data/laptops.csv")
```
# Big Picture
We have all been presented with two columns of quantitative data whether in Excel, R, Python, etc and made a scatterplot and fit a regression line to it. However, because of the simplicity of a a couple of lines of code or selecting a button, it may be easy to overlook how that predicted line was created. By analyzing the concept of residuals we can begin to understand how this predicted line represents the predicted average y value per unit change of x.

## Residual

A residual is the y distance of a given observation or dot from the fitted line. The definition for a given point is $r_i = Y_i - \hat{Y}_i$ with $Y_i$ representing the Y position of the data and $\hat{Y}_i$ representing the Y position of the fitted line.

```{r, message = FALSE, warning = FALSE}
line <- function(x){
  2*x+3
}

p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))+
  geom_line(data = data.frame(x_val = c(0,0), y_val = c(3,0)), aes(x = x_val, y = y_val), group = 1)+ 
  geom_point(data = data.frame(x = 0, y = 3), aes(x,y), size = 2)+
  stat_function(fun = line) + xlim(-5,5)+
  geom_point(data = data.frame(x = 0, y = 0), aes(x,y), size = 2)+
  geom_text(label = "Test", aes(0.6,1))
p
```

