---
title: "Recovering a True Model"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
p1Data <- read_csv("p1Data-1.csv")
p2Data <- read_csv("p2Data-1.csv")
p3Data <- read_csv("p3Data-1.csv")
```

```{r}
pairs(p1Data)
pairs(p1Data, panel = panel.smooth, col = as.factor(p1Data$X2))
```

```{r}
mylm <- lm(Y~X4, data = p1Data)
summary(mylm)

plot(mylm$residuals~., data = p1Data)

mylm2 <- lm(Y~X4 + X2, data = p1Data)
summary(mylm2)

mylm2 <- lm(Y~X2, data = p1Data)
summary(mylm2)
plot(mylm2$residuals~., data = p1Data)

pairs(cbind(Res = mylm2$residuals, p1Data), panel = panel.smooth, col = as.factor)

mylm3 <- lm(Y~X4 + X2 + X4:X2, data = p1Data)
summary(mylm3)

mylm4 <- lm(Y~X4 + X4:X2, data = p1Data)
summary(mylm4)
```


## Part 1 - Recovering a True Model (Beginner Level)

We believe the true model to be...

$$
  Y_i = \underbrace{\beta_0 + \beta_1X_{4i} + \beta_2X{4i}X_{2i}}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = -0.7065 + 2.3958 X_{4i} + 1.7763 X_{4i} X_{2i}
$$

with our estimate of $\sigma$ as 13.11




## Part 2 - Recovering a True Model (Intermediate Level)

```{r}
pairs(p2Data, panel = panel.smooth)
```

```{r}
mylm <- lm(Y~X5 + I(X5^2), data = p2Data)
plot(mylm$residuals~.,data = p2Data)

mylm2 <- lm(Y~X5 + I(X5^2) + X3, data = p2Data)
summary(mylm2)
```

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}

```



## Part 3 - Recovering a True Model (Advanced Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}

```
